---
title: "Calls vs IMD"
author: "NK MA RS"
date: "08/06/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# Load packages
library(fable)
library(lubridate)
library(tsibble)
library(tidyverse)
library(readxl)
library(sf)

# Read and clean data

# excel_sheets("File_1_-_IMD2019_Index_of_Multiple_Deprivation.xlsx")
# imd <- read_xlsx(here::here("File_1_-_IMD2019_Index_of_Multiple_Deprivation.xlsx", sheet = "IMD2019"))  %>%  janitor::clean_names()

# imd <- read_xlsx("C:\\Users\\maczo\\Desktop\\research\\n8_cov\\File_1_-_IMD2019_Index_of_Multiple_Deprivation.xlsx", sheet = "IMD2019") %>%  janitor::clean_names()

imd <- read_xlsx("/Users/reka/Dropbox (The University of Manchester)/N8_covid/analysis/n8_covid/File_1_-_IMD2019_Index_of_Multiple_Deprivation.xlsx", sheet = "IMD2019") %>%  janitor::clean_names()

cheshire_lsoas <- st_read("../cheshire_lsoas.geojson")
cheshire_imds <- left_join(cheshire_lsoas, imd, by = c("LSOA11NM" = "lsoa_name_2011"))

# calls <- read_csv("Z:\\n8_data_v2.csv.gz") %>%  janitor::clean_names()
calls <- read_csv("/Volumes/n8_covid/n8_data_v2.csv.gz") %>%  janitor::clean_names()

minor_categories <- calls %>% 
  count(incident_type) %>% 
  filter(n < 1000) %>% 
  pull(incident_type)

# Categorise calls
calls <- mutate(
  calls,
  incident_type_new = case_when(
    incident_type %in% c("Distraction Burglary") ~
      "Burglary - Residential",
    incident_type %in% c(
      "Assistance to Other Agencies",
      "CRB Use Only",
      "CSI To Be Informed",
      "External System",
      "Found Stolen Vehicle",
      "Message to Pass/OOF Enquiries",
      "PNC Markers",
      "Police Vehicle Recovery",
      "Pre-Planned Events",
      "Stop Search",
      "Training"
    ) ~ "other - admin",
    incident_type %in% c(
      "Alarm - Activation", 
      "Alarm - No Response", 
      "Audible Only Alarm",
      "Insecure Premises",
      "Police Installed Alarm",
      "Unauthorised Encampment"
    ) ~ "alarms",
    incident_type %in% c(
      "Bail Breaches/Wanted Person",
      "Bail/Curfew/Wanted",
      "Prison Licence Recall",
      "Warrant Crown Court"
    ) ~ "warrant/bail issue",
    incident_type %in% c("Domestic Animal Concern", "Wildlife Matters") ~ 
      "animals",
    incident_type %in% c(
      "Firearms - Non Notifiable Crime",
      "Firearms Involved Crime"
    ) ~ "firearms",
    incident_type %in% c("RTC", "RTC - Damage Only") ~ "traffic collision",
    incident_type %in% minor_categories ~ "other - minor",
    TRUE ~ incident_type
  )
)

cheshire_calls <- calls %>% filter(lsoa %in% unique(cheshire_imds$LSOA11NM))

lsoa_filler <- bind_rows(data.frame(lsoa = unique(cheshire_imds$LSOA11NM),
                          pre_post = "Pre-COVID"), 
                         data.frame(lsoa = unique(cheshire_imds$LSOA11NM),
                          pre_post = "2020"))
                         

plot_imd <- function(inc_type) {
  
  type_calls <- cheshire_calls %>% filter(incident_type_new == inc_type) %>% 
    group_by(lsoa) %>% count() %>% 
    left_join(cheshire_imds, ., by = c("LSOA11NM" = "lsoa")) %>% 
    mutate(n = replace_na(n, 0))
  
  labs_df <- type_calls %>% group_by(index_of_multiple_deprivation_imd_decile) %>% count()
  
 p <-  ggplot(type_calls %>% filter(!is.na(index_of_multiple_deprivation_imd_decile)), aes(x = as.factor(index_of_multiple_deprivation_imd_decile), y = n )) + 
    geom_boxplot() + 
    geom_label(data = labs_df, aes(x = index_of_multiple_deprivation_imd_decile, y = 50000, label = paste0("n = ",n))) + 
    scale_y_log10() + 
    labs(title = paste0(inc_type, " calls from LSOAs in each IMD decile 2015-2020")) + 
    xlab("IMD decile (1 = most and 10 = least deprived)") + 
    ylab(paste0("Number of ",inc_type," calls")) + 
    theme_bw()
 
 return(p)
  
}


plot_imd_prepost <- function(inc_type) {
  
  pre_calls <- cheshire_calls %>% 
    filter(incident_type_new == inc_type) %>% 
  mutate(pre_post = ifelse(incident_date_time < dmy("31/01/2020"), "Pre-COVID", "2020"), 
         pre_post = factor(pre_post, levels = c("Pre-COVID", "2020")), 
         inc_wk = yearweek(incident_date_time)) %>%  
    group_by(lsoa, pre_post, inc_wk) %>% 
    count() %>% 
    group_by(lsoa, pre_post) %>% summarise(avg_wkly_calls = mean(n, na.rm=TRUE)) %>% 
  left_join(lsoa_filler, .) %>%
    mutate(avg_wkly_calls = replace_na(avg_wkly_calls, 0)) %>% 
    left_join(., cheshire_imds, by = c("lsoa" = "LSOA11NM")) %>% 
  filter(!is.na(index_of_multiple_deprivation_imd_decile)) %>%  # check why NAs
  mutate(avg_wkly_calls = replace_na(avg_wkly_calls, 0)) 

labs_df <- pre_calls %>% 
  group_by(index_of_multiple_deprivation_imd_decile) %>% summarise(n = n_distinct(lsoa))

 p <-  ggplot() + 
  geom_boxplot(data = pre_calls, aes(x = as.factor(index_of_multiple_deprivation_imd_decile), 
                                     y = avg_wkly_calls, 
                                     color = factor(pre_post, levels =  c("Pre-COVID", "2020")))) + 
  geom_label(data = labs_df, aes(x = index_of_multiple_deprivation_imd_decile, 
                                 y = max(pre_calls$avg_wkly_calls, na.rm = TRUE) + (max(pre_calls$avg_wkly_calls, na.rm = TRUE)/4), 
                                 label = paste0("n = ",n))) +
  labs(title = "Average weekly calls from LSOAs in each IMD decile pre-COVID and 2020") + 
  xlab("IMD decile (1 = most and 10 = least deprived)") + 
  ylab(paste0("Average number of ",inc_type," calls (per week)")) + 
  scale_color_brewer("",type = "qual", palette = 2) +  
  theme_bw() + 
  theme(legend.position = c(0.87, 0.75))
 return(p)
  
}

```


**NOTE** there are 17 LSOAs where the IMD ranking isn't there? It's not in the spreadsheet, IDK why so need to check. They are: `r cheshire_imds %>% filter(is.na(index_of_multiple_deprivation_imd_decile)) %>% pull(LSOA11NM)`


## Total calls by IMD decile

```{r}

total_calls <- cheshire_calls %>% group_by(lsoa) %>% count() %>% 
  left_join(., imd, by = c("lsoa" =  "lsoa_name_2011")) %>%
  filter(!is.na(index_of_multiple_deprivation_imd_decile))  # check why NAs


labs_df <- total_calls  %>% 
  group_by(index_of_multiple_deprivation_imd_decile) %>% count()

ggplot(total_calls , aes(x = as.factor(index_of_multiple_deprivation_imd_decile), y = n )) + 
  geom_boxplot() + 
  geom_label(data = labs_df, aes(x = index_of_multiple_deprivation_imd_decile, y = 50000, label = paste0("n = ",n))) + 
  scale_y_log10() + 
  labs(title = "Total calls from LSOAs in each IMD decile 2015-2020") + 
  xlab("IMD decile (1 = most and 10 = least deprived)") + 
  ylab("Number of calls (all types)") + 
  theme_bw()


```


```{r prepostcovidbox}

pre_calls <- cheshire_calls %>% 
  mutate(pre_post = ifelse(incident_date_time < dmy("31/01/2020"), "Pre-COVID", "2020"), 
         inc_wk = yearweek(incident_date_time)) %>%   
  group_by(lsoa, pre_post, inc_wk) %>% count() %>% 
  group_by(lsoa, pre_post) %>%  summarise(avg_wk_calls = mean(n, na.rm = TRUE)) %>% 
  left_join(lsoa_filler, .) %>%
    mutate(avg_wk_calls = replace_na(avg_wk_calls, 0)) %>% 
  left_join(., cheshire_imds, by = c( "lsoa" = "LSOA11NM")) %>%
  filter(!is.na(index_of_multiple_deprivation_imd_decile)) %>%  # check why NAs
  mutate(avg_wk_calls = replace_na(avg_wk_calls, 0)) 


labs_df <- pre_calls %>% group_by(pre_post, index_of_multiple_deprivation_imd_decile) %>% count()


ggplot() + 
  geom_boxplot(data = pre_calls, aes(x = as.factor(index_of_multiple_deprivation_imd_decile), 
                      y = avg_wk_calls, 
                      color = factor(pre_post, levels =  c("Pre-COVID", "2020")))) + 
  geom_label(data = labs_df, aes(x = index_of_multiple_deprivation_imd_decile, y = max(pre_calls$avg_wk_calls + 20), label = paste0("n = ",n))) +
  labs(title = "Average weekly calls from LSOAs in each IMD decile pre and post COVID") + 
  xlab("IMD decile (1 = most and 10 = least deprived)") + 
  ylab("Average weekly calls") + 
  theme_bw() + 
  scale_color_brewer("",type = "qual", palette = 2) + 
  theme(legend.position = c(0.87, 0.75))

```


## Same but by call type

## Admin

```{r Admin}
plot_imd("Advice/Directions/Admin")
plot_imd_prepost("Advice/Directions/Admin")

```


## Alarms

```{r Alarms}
plot_imd("alarms")
plot_imd_prepost("alarms")

```

## Widlife Concern

```{r Wildlife}
plot_imd("animals")
plot_imd_prepost("animals")
```


## ASB

```{r ASB}
plot_imd("ASB")
plot_imd_prepost("ASB")
```




## Burglary - Commercial

```{r Burglary - Commercial}
plot_imd("Burglary - Business/Commercial")
plot_imd_prepost("Burglary - Business/Commercial")
```



## Burglary - Residential

```{r Burglary - Residential}
plot_imd("Burglary - Residential")
plot_imd_prepost("Burglary - Residential")
```



## Civil Disputes

```{r Civil Disputes}

plot_imd("Civil Disputes")
plot_imd_prepost("Civil Disputes")
```



## Police Safety 

```{r Police Safety}

plot_imd("Code Zero")
plot_imd_prepost("Code Zero")
```


## Minor Injuries 

```{r Minor Injuries}

plot_imd("Collapse/Illness/Injury/Trapped")
plot_imd_prepost("Collapse/Illness/Injury/Trapped")
```

## Complaints Against Police

```{r Compaints Against Police}

plot_imd("Complaints Against Police")
plot_imd_prepost("Complaints Against Police")
```



## Concern for Safety 

```{r Concern For Safety}
plot_imd("Concern For Safety")
plot_imd_prepost("Concern For Safety")
```


## Criminal Damage 

```{r Criminal Damage}
plot_imd("Criminal Damage")
plot_imd_prepost("Criminal Damage")
```



## Domestic Incident

```{r Domestic Incident}
plot_imd("Domestic Incident")
plot_imd_prepost("Domestic Incident")
```




## Drug Related Offences 


```{r Drug Related Offences}
plot_imd("Drugs")
plot_imd_prepost("Drugs")
```




## Firearms 


```{r Firearms}
plot_imd("firearms")
plot_imd_prepost("firearms")
```



## Fraud


```{r Fraud}
plot_imd("Fraud")
plot_imd_prepost("Fraud")
```



## Hate Incident 


```{r Hate Incident}
plot_imd("Hate Incident")
plot_imd_prepost("Hate Incident")
```



## Highway Disruption 

```{r Highway Disruption}
plot_imd("Highway Disruption")
plot_imd_prepost("Highway Disruption")
```




## Missing Person 

```{r Missing Person}
plot_imd("Missing Person")
plot_imd_prepost("Missing Person")
```



## Other - Admin Calls

```{r Other -Admin Calls}
plot_imd("other - admin")
plot_imd_prepost("other - admin")
```


## Other - minor 


```{r Other - minor}
plot_imd("other - minor")
plot_imd_prepost("other - minor")
```


 
## Police Generated 

```{r Police Generated}
plot_imd("Police Generated")
plot_imd_prepost("Police Generated")
```



## Potential Breach of the Peace


```{r Breach of Peace}
plot_imd("Prevent Potential Breach of the Peace")
plot_imd_prepost("Prevent Potential Breach of the Peace")
```



## Property Lost/Found 


```{r Property Lost/Found}
plot_imd("Property Lost/Found")
plot_imd_prepost("Property Lost/Found")
```




## Road Related Traffic Offence

```{r Road Related Traffic}
plot_imd("Road Related Traffic Offence")
plot_imd_prepost("Road Related Traffic Offence")
```


## Robbery 

```{r Robbery}
plot_imd("Robbery")
plot_imd_prepost("Robbery")
```


## Sexual Offences


```{r Sexual Offences}
plot_imd("Sexual Offences")
plot_imd_prepost("Sexual Offences")
```




## Shoplifitng

```{r Shoplifting}
plot_imd("Shoplifting")
plot_imd_prepost("Shoplifting")
```



## Sudden Death 

```{r Sudden Death}
plot_imd("Sudden Death")
plot_imd_prepost("Sudden Death")
```



## Suspicious Activity 

```{r Suspicious Activity}
plot_imd("Suspicious Activity")
plot_imd_prepost("Suspicious Activity")

```



## Suspicious Package 

```{r Suspicious Package}
plot_imd("Suspicious Package")
plot_imd_prepost("Suspicious Package")

```



## Theft from Motor Vehicle

```{r Theft From Motor Vehicle}
plot_imd("Theft From Motor Vehicle")
plot_imd_prepost("Theft From Motor Vehicle")

```



## Theft of Motor Vehicle 


```{r Theft of Motor Vehicle}
plot_imd("Theft Of Motor Vehicle")
plot_imd_prepost("Theft Of Motor Vehicle")

```



## Theft Other 

```{r Theft Other}
plot_imd("Theft Other")
  plot_imd_prepost("Theft Other")

```



## Traffic collision

```{r Traffic Collisions}
plot_imd("traffic collision")
plot_imd_prepost("traffic collision")

```



## Violence and Harassment Incidents 

```{r Violence/Harassment}
plot_imd("Violence/Harassment")
plot_imd_prepost("Violence/Harassment")

```



## Warrant and Bail Incidents 


```{r Warrant/Bail}
plot_imd("warrant/bail issue")
plot_imd_prepost("warrant/bail issue")
```


# What about change over time for each IMD


```{r}



weekly_count_by_lsoa <- cheshire_calls %>% 
   mutate(inc_wk = as_date(yearweek(incident_date_time))) %>%   
   group_by(lsoa, inc_wk) %>% count() %>% 
  left_join(lsoa_filler, .) %>%
    mutate(n = replace_na(n, 0)) %>% 
    left_join(., cheshire_imds, by = c("lsoa" = "LSOA11NM")) %>% 
  filter(!is.na(index_of_multiple_deprivation_imd_decile)) # check why NAs

firstweek <- min(weekly_count_by_lsoa$inc_wk)
lastweek <- max(weekly_count_by_lsoa$inc_wk)


weekly_count_by_lsoa <- weekly_count_by_lsoa %>% 
  filter(inc_wk > firstweek & inc_wk < lastweek) 

num_lsoa_per_decile <- weekly_count_by_lsoa %>% 
  select(lsoa, index_of_multiple_deprivation_imd_decile) %>% 
  distinct() %>% 
  group_by(index_of_multiple_deprivation_imd_decile) %>% 
  count() %>% 
  rename(num_lsoas = n)

thing <- weekly_count_by_lsoa %>% 
  left_join(., num_lsoa_per_decile, by = c("index_of_multiple_deprivation_imd_decile" = "index_of_multiple_deprivation_imd_decile")) %>% 
  select(inc_wk, n, index_of_multiple_deprivation_imd_decile, num_lsoas)  %>% 
  group_by(inc_wk, index_of_multiple_deprivation_imd_decile, num_lsoas) %>%
  summarise(n = sum(n)) %>%
  mutate(calls_per_lsoa = n/num_lsoas) %>% 
  group_by(inc_wk, index_of_multiple_deprivation_imd_decile) %>%
  summarise(calls_per_lsoa = calls_per_lsoa) %>% 
    mutate(percentage = calls_per_lsoa / sum(calls_per_lsoa))

# Plot
ggplot(thing, aes(x=inc_wk, y=percentage, fill=as.factor(index_of_multiple_deprivation_imd_decile))) + 
    geom_area(alpha= 1 , size=0.001, colour="black") +
  # First UK case
  geom_vline(xintercept = as.Date("2020-01-23"), linetype = "33") +
  # First UK lockdown begins
  geom_vline(xintercept = as.Date("2020-03-23"), linetype = "12") +
  # First UK lockdown ends
  geom_vline(xintercept = as.Date("2020-06-01"), linetype = "42") +
  scale_x_date(date_breaks = "6 months", date_labels = "%b\n%Y", expand = c(0, 0)) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_brewer(type = "qual", palette = "Paired") + 
  labs(
    title = "Proportion of Calls from different IMD deciles",
    x = NULL,
    y = "Proportion",
    fill = "IMD decile"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") +    guides(fill=guide_legend(nrow=1,byrow=TRUE))


```


zoom 

```{r}



# Plot
ggplot(thing %>% filter(inc_wk > as.Date("2019-10-10")), 
       aes(x=inc_wk, y=percentage, fill=as.factor(index_of_multiple_deprivation_imd_decile))) + 
    geom_area(alpha= 1 , size=0.001, colour="black") +
  # First UK case
  geom_vline(xintercept = as.Date("2020-01-23"), linetype = "33") +
  # First UK lockdown begins
  geom_vline(xintercept = as.Date("2020-03-23"), linetype = "12") +
  # First UK lockdown ends
  geom_vline(xintercept = as.Date("2020-06-01"), linetype = "42") +
  scale_x_date(date_breaks = "6 months", date_labels = "%b\n%Y", expand = c(0, 0)) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_brewer(type = "qual", palette = "Paired") + 
  labs(
    title = "Proportion of Calls from different IMD deciles",
    x = NULL,
    y = "Proportion",
    fill = "IMD decile"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") +    guides(fill=guide_legend(nrow=1,byrow=TRUE))


```



## Breakdown by call type

```{r}


imd_area_plot <- function(inc_type) {
  
  type_calls <- cheshire_calls %>% filter(incident_type_new == inc_type) %>% 
    group_by(lsoa) %>% count() %>% 
    left_join(cheshire_imds, ., by = c("LSOA11NM" = "lsoa")) %>% 
    mutate(n = replace_na(n, 0))
  
  labs_df <- type_calls %>% group_by(index_of_multiple_deprivation_imd_decile) %>% count()
  
 p <-  ggplot(type_calls %>% filter(!is.na(index_of_multiple_deprivation_imd_decile)), aes(x = as.factor(index_of_multiple_deprivation_imd_decile), y = n )) + 
    geom_boxplot() + 
    geom_label(data = labs_df, aes(x = index_of_multiple_deprivation_imd_decile, y = 50000, label = paste0("n = ",n))) + 
    scale_y_log10() + 
    labs(title = paste0(inc_type, " calls from LSOAs in each IMD decile 2015-2020")) + 
    xlab("IMD decile (1 = most and 10 = least deprived)") + 
    ylab(paste0("Number of ",inc_type," calls")) + 
    theme_bw()
 
 return(p)
  
}


# Plot
ggplot(thing %>% filter(inc_wk > as.Date("2019-10-10")), 
       aes(x=inc_wk, y=percentage, fill=as.factor(index_of_multiple_deprivation_imd_decile))) + 
    geom_area(alpha= 1 , size=0.001, colour="black") +
  # First UK case
  geom_vline(xintercept = as.Date("2020-01-23"), linetype = "33") +
  # First UK lockdown begins
  geom_vline(xintercept = as.Date("2020-03-23"), linetype = "12") +
  # First UK lockdown ends
  geom_vline(xintercept = as.Date("2020-06-01"), linetype = "42") +
  scale_x_date(date_breaks = "6 months", date_labels = "%b\n%Y", expand = c(0, 0)) +
  scale_y_continuous(labels = scales::percent_format()) +
  scale_fill_brewer(type = "qual", palette = "Paired") + 
  labs(
    title = "Proportion of Calls from different IMD deciles",
    x = NULL,
    y = "Proportion",
    fill = "IMD decile"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom") +    guides(fill=guide_legend(nrow=1,byrow=TRUE))


```
