---
title: "Calls vs IMD"
author: "NK MA RS"
date: "08/06/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

# Load packages
library(fable)
library(lubridate)
library(tsibble)
library(tidyverse)
library(readxl)

# Read and clean data

# excel_sheets("File_1_-_IMD2019_Index_of_Multiple_Deprivation.xlsx")
# imd <- read_xlsx(here::here("File_1_-_IMD2019_Index_of_Multiple_Deprivation.xlsx", sheet = "IMD2019"))  %>%  janitor::clean_names()

imd <- read_xlsx("C:\\Users\\maczo\\Desktop\\research\\n8_cov\\File_1_-_IMD2019_Index_of_Multiple_Deprivation.xlsx", sheet = "IMD2019") %>%  janitor::clean_names()

calls <- read_csv("Z:\\n8_data_v2.csv.gz") %>%  janitor::clean_names()

minor_categories <- calls %>% 
  count(incident_type) %>% 
  filter(n < 1000) %>% 
  pull(incident_type)

# Categorise calls
calls <- mutate(
  calls,
  incident_type_new = case_when(
    incident_type %in% c("Distraction Burglary") ~
      "Burglary - Residential",
    incident_type %in% c(
      "Assistance to Other Agencies",
      "CRB Use Only",
      "CSI To Be Informed",
      "External System",
      "Found Stolen Vehicle",
      "Message to Pass/OOF Enquiries",
      "PNC Markers",
      "Police Vehicle Recovery",
      "Pre-Planned Events",
      "Stop Search",
      "Training"
    ) ~ "other - admin",
    incident_type %in% c(
      "Alarm - Activation", 
      "Alarm - No Response", 
      "Audible Only Alarm",
      "Insecure Premises",
      "Police Installed Alarm",
      "Unauthorised Encampment"
    ) ~ "alarms",
    incident_type %in% c(
      "Bail Breaches/Wanted Person",
      "Bail/Curfew/Wanted",
      "Prison Licence Recall",
      "Warrant Crown Court"
    ) ~ "warrant/bail issue",
    incident_type %in% c("Domestic Animal Concern", "Wildlife Matters") ~ 
      "animals",
    incident_type %in% c(
      "Firearms - Non Notifiable Crime",
      "Firearms Involved Crime"
    ) ~ "firearms",
    incident_type %in% c("RTC", "RTC - Damage Only") ~ "traffic collision",
    incident_type %in% minor_categories ~ "other - minor",
    TRUE ~ incident_type
  )
)

plot_imd <- function(inc_type) {
  
  type_df <- calls %>% filter(incident_type_new == inc_type)
  
  type_calls <- type_df %>% group_by(lsoa) %>% count() %>% 
    left_join(., imd, by = c("lsoa" =  "lsoa_name_2011")) 
  
  labs_df <- type_calls %>% group_by(index_of_multiple_deprivation_imd_decile) %>% count()
  
 p <-  ggplot(type_calls %>% filter(!is.na(index_of_multiple_deprivation_imd_decile)), aes(x = as.factor(index_of_multiple_deprivation_imd_decile), y = n )) + 
    geom_boxplot() + 
    geom_label(data = labs_df, aes(x = index_of_multiple_deprivation_imd_decile, y = 50000, label = paste0("n = ",n))) + 
    scale_y_log10() + 
    labs(title = paste0(inc_type, " calls from LSOAs in each IMD decile 2015-2020")) + 
    xlab("IMD decile (1 = most and 10 = least deprived)") + 
    ylab(paste0("Number of ",inc_type," calls")) + 
    theme_bw()
 
 return(p)
  
}

```


## Total calls by IMD decile

```{r}

total_calls <- calls %>% group_by(lsoa) %>% count() %>% 
  left_join(., imd, by = c("lsoa" =  "lsoa_name_2011")) 

labs_df <- total_calls %>% group_by(index_of_multiple_deprivation_imd_decile) %>% count()

ggplot(total_calls %>% filter(!is.na(index_of_multiple_deprivation_imd_decile)), aes(x = as.factor(index_of_multiple_deprivation_imd_decile), y = n )) + 
  geom_boxplot() + 
  geom_label(data = labs_df, aes(x = index_of_multiple_deprivation_imd_decile, y = 50000, label = paste0("n = ",n))) + 
  scale_y_log10() + 
  labs(title = "Total calls from LSOAs in each IMD decile 2015-2020") + 
  xlab("IMD decile (1 = most and 10 = least deprived)") + 
  ylab("Number of calls (all types)") + 
  theme_bw()


```


## Same but by call type

## Admin

```{r Admin}
plot_imd("Advice/Directions/Admin")
```


## Alarms

```{r Alarms}
plot_imd("alarms")
```

## Widlife Concern

```{r Wildlife}
plot_imd("animals")
```


## ASB

```{r ASB}
plot_imd("ASB")
```




## Burglary - Commercial

```{r Burglary - Commercial}
plot_imd("Burglary - Business/Commercial")
```



## Burglary - Residential

```{r Burglary - Residential}
plot_imd("Burglary - Residential")
```



## Civil Disputes

```{r Civil Disputes}

plot_imd("Civil Disputes")
```



## Police Safety 

```{r Police Safety}

plot_imd("Code Zero")
```


## Minor Injuries 

```{r Minor Injuries}

plot_imd("Collapse/Illness/Injury/Trapped")
```

## Complaints Against Police

```{r Compaints Against Police}

plot_imd("Complaints Against Police")
```



## Concern for Safety 

```{r Concern For Safety}
plot_imd("Concern For Safety")
```


## Criminal Damage 

```{r Criminal Damage}
plot_imd("Criminal Damage")
```



## Domestic Incident

```{r Domestic Incident}
plot_imd("Domestic Incident")
```




## Drug Related Offences 


```{r Drug Related Offences}
plot_imd("Drugs")
```




## Firearms 


```{r Firearms}
plot_imd("firearms")
```



## Fraud


```{r Fraud}
plot_imd("Fraud")
```



## Hate Incident 


```{r Hate Incident}
plot_imd("Hate Incident")
```



## Highway Disruption 

```{r Highway Disruption}
plot_imd("Highway Disruption")
```




## Missing Person 

```{r Missing Person}
plot_imd("Missing Person")
```



## Other - Admin Calls

```{r Other -Admin Calls}
plot_imd("other - admin")
```


## Other - minor 


```{r Other - minor}
plot_imd("other - minor")
```


 
## Police Generated 

```{r Police Generated}
plot_imd("Police Generated")
```



## Potential Breach of the Peace


```{r Breach of Peace}
plot_imd("Prevent Potential Breach of the Peace")
```



## Property Lost/Found 


```{r Property Lost/Found}
plot_imd("Property Lost/Found")
```




## Road Related Traffic Offence

```{r Road Related Traffic}
plot_imd("Road Related Traffic Offence")
```


## Robbery 

```{r Robbery}
plot_imd("Robbery")
```


## Sexual Offences


```{r Sexual Offences}
plot_imd("Sexual Offences")
```




## Shoplifitng

```{r Shoplifting}
plot_imd("Shoplifting")
```



## Sudden Death 

```{r Sudden Death}
plot_imd("Sudden Death")
```



## Suspicious Activity 

```{r Suspicious Activity}
plot_imd("Suspicious Activity")
```



## Suspicious Package 

```{r Suspicious Package}
plot_imd("Suspicious Package")
```



## Theft from Motor Vehicle

```{r Theft From Motor Vehicle}
plot_imd("Theft From Motor Vehicle")
```



## Theft of Motor Vehicle 


```{r Theft of Motor Vehicle}
plot_imd("Theft Of Motor Vehicle")
```



## Theft Other 

```{r Theft Other}
plot_imd("Theft Other")
```



## Traffic collision

```{r Traffic Collisions}
plot_imd("traffic collision")
```



## Violence and Harassment Incidents 

```{r Violence/Harassment}
plot_imd("Violence/Harassment")
```



## Warrant and Bail Incidents 


```{r Warrant/Bail}
plot_imd("warrant/bail issue")
```







