---
title: "Calls for Service During 2021 Compared to Pre-Pandemic Forecast - Incident Type"
author: "Matt Ashby, Nadia Kennar, Reka Solymosi"
date: "23/04/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(cache = TRUE, echo = FALSE, include = TRUE)

library(lubridate)
library(tidyverse)
library(tsibble)
```

```{r load data, include=FALSE}
call_forecasts <- read_rds(here::here("output/call_forecasts.Rds"))
#call_forecasts <- read_rds("Z:\\call_forecasts.Rds")

# Note important dates
# Source: https://www.instituteforgovernment.org.uk/sites/default/files/timeline-lockdown-web.pdf
dates <- tribble(
  ~date, ~event,
  "2020-01-31", "first UK COVID case",
  "2020-03-23", "first lockdown begins",
  "2020-06-15", "first lockdown ends",
  "2020-11-05", "second lockdown begins",
  "2020-12-02", "second lockdown ends"
) %>% 
  mutate(
    date = as_date(yearweek(ymd(date))), 
    row = row_number(),
    label = str_glue("{row}. {event}")
  )
```

```{r create chart function, include=FALSE}
forecast_chart <- function (forecasts, types) {
  
  forecasts %>% 
    filter(incident_type_new %in% types) %>% 
    # Occasionally the ARIMA model may produce forecast confidence intervals
    # that are less than zero, which will not show up on the plot because we
    # have set the y axis to start at zero (to make changes in vertical position
    # on the axis representative of changes in magnitude). To deal with this we
    # will manually set the lower CI of the forecasts to be zero if it is less
    # than zero in the original forecast.
    mutate(forecast_lower = ifelse(forecast_lower < 0, 0, forecast_lower)) %>% 
    ggplot() +
    # Forecast
    geom_ribbon(
      aes(incident_week, ymin = forecast_lower, ymax = forecast_upper), 
      na.rm = TRUE,
      alpha = 0.5, 
      fill = "grey80"
    ) +
    geom_line(aes(incident_week, forecast_mean), na.rm = TRUE, linetype = "22") +
    # Dates of interest
    geom_vline(aes(xintercept = date), data = dates, linetype = "12") +
    geom_label(aes(date, 0, label = row), data = dates, colour = "grey20") +
    # Actual calls
    geom_line(aes(incident_week, actual_calls)) +
    geom_point(aes(incident_week, actual_calls, fill = sig), shape = 21) +
    scale_x_date(date_labels = "%e %b\n%Y", 
                 limits = as.Date(c("2020-01-06", "2020-12-21"))) +
    scale_y_continuous(limits = c(0, NA), labels = scales::comma_format()) +
    scale_fill_manual(values = c(`TRUE` = "black", `FALSE` = "grey80")) +
    labs(
      title = "Calls for service during 2021 compared to pre-pandemic forecast",
      subtitle = str_wrap(
        str_glue("Events by week: ", str_c(pull(dates, label), collapse = "; ")), 
        80
      ),
      caption = "Forecast calculated using data up to 31 January 2020",
      x = NULL,
      y = "weekly count of calls for service",
      fill = "actual calls significantly different from forecast"
    ) +
    theme_minimal() +
    theme(legend.position = "bottom", plot.title = element_text(face = "bold"))
  
}
```



## Abondended Phone Call

```{r Abandoned}
forecast_chart(call_forecasts, "Abandoned Phone Call")
```

Majority of calls are significantly different from the forecast data. Less weekly calls received with a decreasing trend 





## Admin

```{r Admin}
forecast_chart(call_forecasts, "Advice/Directions/Admin")
```

No significant differences in the weekly call counts from Admin. Nine time periods fall above of the conifdence intervals, mainly following the first UK lockdown


## Alarms

```{r Alarms}
forecast_chart(call_forecasts, "alarms")
```
No significant differences between actual calls and forecast data


## Widlife Concern

```{r Wildlife}
forecast_chart(call_forecasts, "animals")
```

No significant differences between actual calls and forecast data


## ASB

```{r ASB}
forecast_chart(call_forecasts, "ASB")
```


Over 20 point are considered significantly different form the forecast data as expected, mainly following the first lockdown and second lockdown. Supports literature with increase in ASB due to breaking restriction rules. Interesting to note the reduction of points outside the confidence intervals from the first lockdown to the second lockdown. 



## Burglary - Commercial

```{r Burglary - Commercial}
forecast_chart(call_forecasts, "Burglary - Business/Commercial")
```

General decrease of commercial calls with little significant difference following the ending of first lockdown


## Burglary - Residential

```{r Burglary - Residential}
forecast_chart(call_forecasts, "Burglary - Residential")
```

Majority of the calls are significantly different form the forecast, with a sharper reduction at the end of the second lockdown period. However, to say there is a clear difference in the trend is questionable


## Civil Disputes

```{r Civil Disputes}

forecast_chart(call_forecasts, "Civil Disputes")
```

Actual calls are general lower than the forecast data but nothing too noticeable as only one point is significantly different from the forecast


## Police Safety 

```{r Police Safety}

forecast_chart(call_forecasts, "Code Zero")
```

No significant differences in the trend between actual calls and foretasted data. A few points do sit above and below but this is expected with calls to service


## Minor Injuries 

```{r Minor Injuries}

forecast_chart(call_forecasts, "Collapse/Illness/Injury/Trapped")
```

Foretasted trend seems to follow actual trend, nothing too different to note  



## Complaints Against Police

```{r Compaints Against Police}

forecast_chart(call_forecasts, "Complaints Against Police")
```

Actual calls tend do not differ too much from the forecast data. With only, 5 points below the confidence intervals before the first lockdown ends and then 2 points higher than the confidence intervals after the second lockdown begins 



## Concern for Safety 

```{r Concern For Safety}
forecast_chart(call_forecasts, "Concern For Safety")
```


Actual calls from concerns of safety rarely fall below the lower 95% CIs; no significant differences in the plot 



## Criminal Damage 

```{r Criminal Damage}
forecast_chart(call_forecasts, "Criminal Damage")
```


Following the first lockdown, actual calls of criminal damage are significantly different from the forecast falling below the lower 95% CIs.



## Domestic Incident

```{r Domestic Incident}
forecast_chart(call_forecasts, "Domestic Incident")
```

Only a few of the actual calls recorded for domestic incident fall above the upper bound of the 95% CI, mainly following the start of the first lockdown. Might have expected to see more calls falling outside the predicted data with literature stating that more people remain home?




## Drug Related Offences 


```{r Drug Related Offences}
forecast_chart(call_forecasts,"Drugs")
```


Small change in actual calls received during the start of the first lockdown compared to forecast data. During the start of the first UK lockdown there is a significant increase in drug related offenses. 


## Firearms 


```{r Firearms}
forecast_chart(call_forecasts, "firearms")
```

The overall patterns of actual crimes do not differ that significantly from the forecast results with the majority of points remaining within the upper and lower bounds of the confidence intervals.


## Fraud


```{r Fraud}
forecast_chart(call_forecasts,"Fraud")
```

9 time points are seen as significantly different from the forecast, happening from the first UK Covid case up until the ending of the first lockdown. Most of these points fall below the lower bound of the CIs, could be reflective of less opportunity for Fraud to happen given the closures of retails and businesses. However, with inconsistent changes in the trend a clear conclusion can't be drawn 


## Hate Incident 


```{r Hate Incident}
forecast_chart(call_forecasts,"Hate Incident")
```

Overall actual calls do not significant differ from forecasted data. During the the first UK Covid case and the first lockdown, there are two points above the forecast data. There is also a general increase of hate incidents (within the CIs but still high) following the fist lockdown ending - reflective of increased biased and prejudice to certain communities



## Highway Disruption 

```{r Highway Disruption}
forecast_chart(call_forecasts,"Highway Disruption")
```

Significantly lower levels of highway disruption from the first lockdown as a result of less people traveling. Also supported by the small increase after the first lockdown ended (3)




## Missing Person 

```{r Missing Person}
forecast_chart(call_forecasts, "Missing Person")
```

Significant differences again following the first lockdown. Supported by literature with more children being at home and less opportunity for 'juvenile runaways'.


## Other - Admin Calls

```{r Other -Admin Calls}
forecast_chart(call_forecasts,"other - admin")
```

Small changes to admin calls following the first lockdown as expected with a decrease of weekly calls. However, typically the trends follows the overall forecast pattern


## Other - minor 


```{r Other - minor}
forecast_chart(call_forecasts, "other - minor")
```


Actual calls generally fall within the CIs - no siginificant differences 

 
## Police Generated 

```{r Police Generated}
forecast_chart(call_forecasts, "Police Generated")
```

Interestingly, actual calls from police generated calls fall above the confidence interval throughout all 5 key dates. However, no clear trend in the changes instead some random fluctuations



## Potential Breach of the Peace


```{r Breach of Peace}
forecast_chart(call_forecasts,"Prevent Potential Breach of the Peace")
```

Actual calls of potential breach of peace are much lower than the forecasted data. This makes sense due to changes in peoples daily routine resulting in less opportunity to causes alarm and threat to the community 



## Property Lost/Found 


```{r Property Lost/Found}
forecast_chart(call_forecasts,"Property Lost/Found")
```


No obvious diferences between actual calls and forecasted data.




## Road Related Traffic Offence

```{r Road Related Traffic}
forecast_chart(call_forecasts,"Road Related Traffic Offence")
```

Actual calls of road related traffic offenses typically fall below the forecasted data. Again, this is expected due to less people traveling during the restrictions put in place and a reduction in able employees. 

## Robbery 

```{r Robbery}
forecast_chart(call_forecasts, "Robbery")
```


During time period one, there was a high spike in actual calls of robbery where the first UK Covid case happened - why is this? We also see a break in the seasonality and more random fluctuations throughout the five key dates. Arguably nothing too different from the forecasted data


## Sexual Offences


```{r Sexual Offences}
forecast_chart(call_forecasts, "Sexual Offences")
```


During the first lockdown cases of sexual offenses decreased and fell below the 95% confidence intervals, trends however return to normal following the end of the lockdown. No signifcant differences from the forecast data




## Shoplifitng

```{r Shoplifting}
forecast_chart(call_forecasts, "Shoplifting")
```

Significant differences following the first lockdown as would be expected due to closure of retail and businesses. Arguably different seasonality patterns as well 


## Sudden Death 

```{r Sudden Death}
forecast_chart(call_forecasts, "Sudden Death")
```

Two time periods are seen as significantly different from the forecast which happens at the start of the first UK lockdown, this could be as a result of increase Covid-19 cases or reduced information around the virus, Results tend to fall back to normal soon after. Overall, actual calls are not significantly different from forecast



## Suspicious Activity 

```{r Suspicious Activity}
forecast_chart(call_forecasts, "Suspicious Activity")
```

No significant differences in suspicious activity calls.



## Suspicious Package 

```{r Suspicious Package}
forecast_chart(call_forecasts, "Suspicious Package")
```

Arguable differences in actual calls compared to forecast data.


## Theft from Motor Vehicle

```{r Theft From Motor Vehicle}
forecast_chart(call_forecasts, "Theft From Motor Vehicle")
```

General decrease in theft from motor vehicle following the first UK lockdown. This is as expected with more people working from home and possibly increased home security.



## Theft of Motor Vehicle 


```{r Theft of Motor Vehicle}
forecast_chart(call_forecasts, "Theft Of Motor Vehicle")
```

General decrease of motor theft following the first uk lockdown. However, trends seem to follow the same seasonal pattern just at a lower weekly rate.


## Theft Other 

```{r Theft Other}
forecast_chart(call_forecasts, "Theft Other")
```

General decrease of theft other following the first uk lockdown. Overall, the trend from actual calls are significantly different from forecast data


## Traffic collision

```{r Traffic Collisions}
forecast_chart(call_forecasts, "traffic collision")
```

General decrease following the first lockdown, trends seems to increase soon after but still can be considered significantly different from the forecast data a a decreased weekly rate 


## Violence and Harassment Incidents 

```{r Violence/Harassment}
forecast_chart(call_forecasts, "Violence/Harassment")
```

Following the first lockdown there is a general decrease in weekly counts of violence and harassment compared to the forecast data.


## Warrant and Bail Incidents 


```{r Warrant/Bail}
forecast_chart(call_forecasts, "warrant/bail issue")
```


Nothing too significant to note here, general decrease of calls compared to the forecast data as expected.






Concluding Thoughts; Most significant changes in calls to service happen following the first U.K lockdown, these patterns of change are also reflected in the second lockdown but at a much lower weekly crime rate. 







