---
title: "Response Time"
author: "NK MA RS"
date: "07/06/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
library(lubridate)
library(tsibble)
library(tidyverse)

calls <- read_csv("Z:\\n8_data_v2.csv.gz") %>%  janitor::clean_names()

all_rt_forecast <- read_rds(here::here("output/all_resp_time_forecast.Rds"))
by_type_rt_forecast <- read_rds(here::here("output//by_type_rt_forecasts.Rds"))

# Find categories with fewer than a certain number of calls so we can exclude
# these, since these call types are not amenable to these types of incidents
minor_categories <- calls %>% 
  count(incident_type) %>% 
  filter(n < 1000) %>% 
  pull(incident_type)

# Categorise calls
calls <- mutate(
  calls,
  incident_type_new = case_when(
    incident_type %in% c("Distraction Burglary") ~
      "Burglary - Residential",
    incident_type %in% c(
      "Assistance to Other Agencies",
      "CRB Use Only",
      "CSI To Be Informed",
      "External System",
      "Found Stolen Vehicle",
      "Message to Pass/OOF Enquiries",
      "PNC Markers",
      "Police Vehicle Recovery",
      "Pre-Planned Events",
      "Stop Search",
      "Training"
    ) ~ "other - admin",
    incident_type %in% c(
      "Alarm - Activation", 
      "Alarm - No Response", 
      "Audible Only Alarm",
      "Insecure Premises",
      "Police Installed Alarm",
      "Unauthorised Encampment"
    ) ~ "alarms",
    incident_type %in% c(
      "Bail Breaches/Wanted Person",
      "Bail/Curfew/Wanted",
      "Prison Licence Recall",
      "Warrant Crown Court"
    ) ~ "warrant/bail issue",
    incident_type %in% c("Domestic Animal Concern", "Wildlife Matters") ~ 
      "animals",
    incident_type %in% c(
      "Firearms - Non Notifiable Crime",
      "Firearms Involved Crime"
    ) ~ "firearms",
    incident_type %in% c("RTC", "RTC - Damage Only") ~ "traffic collision",
    incident_type %in% minor_categories ~ "other - minor",
    TRUE ~ incident_type
  )
)




# Note important dates
# Source: https://www.instituteforgovernment.org.uk/sites/default/files/timeline-lockdown-web.pdf
dates <- tribble(
  ~date, ~event,
  "2020-01-31", "first UK COVID case",
  "2020-03-23", "first lockdown begins",
  "2020-06-15", "first lockdown ends",
  "2020-11-05", "second lockdown begins",
  "2020-12-02", "second lockdown ends"
) %>% 
  mutate(
    date = as_date(yearweek(ymd(date))), 
    row = row_number(),
    label = str_glue("{row}. {event}")
  )



forecast_chart <- function (forecasts, types) {
  
  forecasts %>% 
    filter(incident_type_new %in% types) %>% 
    # Occasionally the ARIMA model may produce forecast confidence intervals
    # that are less than zero, which will not show up on the plot because we
    # have set the y axis to start at zero (to make changes in vertical position
    # on the axis representative of changes in magnitude). To deal with this we
    # will manually set the lower CI of the forecasts to be zero if it is less
    # than zero in the original forecast.
    mutate(forecast_lower = ifelse(forecast_lower < 0, 0, forecast_lower)) %>% 
    ggplot() +
    # Forecast
    geom_ribbon(
      aes(incident_week, ymin = forecast_lower, ymax = forecast_upper), 
      na.rm = TRUE,
      alpha = 0.5, 
      fill = "grey80"
    ) +
    geom_line(aes(incident_week, forecast_mean), na.rm = TRUE, linetype = "22") +
    # Dates of interest
    geom_vline(aes(xintercept = date), data = dates, linetype = "12") +
    geom_label(aes(date, 0, label = row), data = dates, colour = "grey20") +
    # Actual calls
    geom_line(aes(incident_week, actual_rt)) +
    geom_point(aes(incident_week, actual_rt, fill = sig), shape = 21) +
    scale_x_date(date_labels = "%e %b\n%Y", 
                 limits = as.Date(c("2020-01-06", "2020-12-21")))+
    scale_y_continuous(limits = c(0, NA), labels = scales::comma_format()) +
    scale_fill_manual(values = c(`TRUE` = "black", `FALSE` = "grey80")) +
    labs(
      title = "Median response time for all calls for service during 2021 compared to pre-pandemic forecast",
      subtitle = str_wrap(
        str_glue("Response time by week: ", str_c(pull(dates, label), collapse = "; ")), 
        80
      ),
      caption = "Forecast calculated using data up to 31 January 2020",
      x = NULL,
      y = "Median response time (in minutes)",
      fill = "Actual response time significantly different from forecast"
    ) +
    theme_minimal() +
    theme(legend.position = "bottom", plot.title = element_text(face = "bold"))
  
}

```

## Overall response time 

```{r}

calls$attended <- case_when(calls$attended_flag == 1 ~ "Yes", 
                            calls$attended_flag == 0 ~ "No", 
                            TRUE ~ NA_character_)

# filter attended only 

attended_calls <- calls %>% filter(attended == "Yes")

# create response time variable

attended_calls$response_time <- as.period(ymd_hms(attended_calls$earliest_arrived_date_time) -  
                                     ymd_hms(attended_calls$earliest_deployed_date_time))

attended_calls$response_time_mins <- time_length(attended_calls$response_time, unit = "minute")


```

Overall there is a wide variation in response time for all calls over the whole data set. The average is `r mean(time_length(attended_calls$response_time, unit = "minute"), na.rm = TRUE)` minutes, the median `r median(time_length(attended_calls$response_time, unit = "minute"), na.rm = TRUE)` minutes, the maximum `r max(time_length(attended_calls$response_time, unit = "minute"), na.rm = TRUE)/60/24` days, and then there are some negative numbers in there (the shortest response time being `r min(time_length(attended_calls$response_time, unit = "minute"), na.rm = TRUE)` minutes). This is just under `r 12*60`, minutes which is 12 hours - so we can assume what is happening is the selection of the wrong time out by 12 hours. To fix this, add 720 to all negative time results.

```{r}

attended_calls$response_time_mins <- ifelse(attended_calls$response_time_mins <0, 
                                            attended_calls$response_time_mins + 720, 
                                            attended_calls$response_time_mins)

options(scipen = 999) # remove scientific notation

resp_plot <- ggplot(attended_calls, aes(x =response_time_mins)) + 
  geom_density(fill = "grey", alpha = 0.7) + 
     geom_vline(xintercept = 1, linetype = 10) + 
  annotate("label", x = 1, y = 0.6, label = "1 minute") + 
   geom_vline(xintercept = 60, linetype = 11) + 
  annotate("label", x = 60, y = 0.6, label = "1 hour") + 
  geom_vline(xintercept = 24*60, linetype = 12) + 
  annotate("label", x = 24*60, y = 0.6, label = "1 day") + 
    geom_vline(xintercept = 24*60*7, linetype = 13) + 
  annotate("label", x = 24*60*7, y = 0.6, label = "1 week") + 
  theme_bw() + 
  scale_x_log10() + 
  labs(title = "Distribution of response time across all calls")+ 
  xlab("Response time in minutes") + 
  theme(axis.text.y = element_blank(), 
        axis.ticks.y = element_blank())

resp_plot

```

How does median response time change over the years? 


```{r}
attended_calls %>% 
  mutate(incident_week = yearweek(incident_date_time)) %>% 
  group_by(incident_week) %>% 
  summarise(median_resp_time = median(response_time_mins, na.rm = TRUE)) %>% 
  ggplot(., aes(x = incident_week, y = median_resp_time)) + 
  geom_point() + 
  geom_line() + 
  theme_bw() + 
  geom_smooth() + 
  xlab("") + 
  ylab("Median response time in minutes")

```


## Response time over the pandemic

What about overall response time vs forecast: 

```{r}

all_rt_forecast %>% 
    mutate(forecast_lower = ifelse(forecast_lower < 0, 0, forecast_lower)) %>% 
    ggplot() +
    # Forecast
    geom_ribbon(
      aes(incident_week, ymin = forecast_lower, ymax = forecast_upper), 
      na.rm = TRUE,
      alpha = 0.5, 
      fill = "grey80"
    ) +
    geom_line(aes(incident_week, forecast_mean), na.rm = TRUE, linetype = "22") +
    # Dates of interest
    geom_vline(aes(xintercept = date), data = dates, linetype = "12") +
    geom_label(aes(date, 0, label = row), data = dates, colour = "grey20") +
    # Actual calls
    geom_line(aes(incident_week, actual_rt)) +
    geom_point(aes(incident_week, actual_rt, fill = sig), shape = 21) +
    scale_x_date(date_labels = "%e %b\n%Y") +
    scale_y_continuous(limits = c(0, NA), labels = scales::comma_format()) +
    scale_fill_manual(values = c(`TRUE` = "black", `FALSE` = "grey80")) +
    labs(
      title = "Median response time for all calls for service during 2021 compared to pre-pandemic forecast",
      caption = "Forecast calculated using data up to 31 January 2020",
      x = NULL,
      y = "Median response time in minutes",
      fill = "Response time significantly different from forecast"
    ) +
    theme_minimal() +
    theme(legend.position = "bottom", plot.title = element_text(face = "bold"))

```
## by call type





Now we can break this down by call type:

## Admin

```{r Admin}
forecast_chart(by_type_rt_forecast, "Advice/Directions/Admin")
```


## Alarms

```{r Alarms}
forecast_chart(by_type_rt_forecast, "alarms")
```

## Widlife Concern

```{r Wildlife}
forecast_chart(by_type_rt_forecast, "animals")
```


## ASB

```{r ASB}
forecast_chart(by_type_rt_forecast, "ASB")
```




## Burglary - Commercial

```{r Burglary - Commercial}
forecast_chart(by_type_rt_forecast, "Burglary - Business/Commercial")
```



## Burglary - Residential

```{r Burglary - Residential}
forecast_chart(by_type_rt_forecast, "Burglary - Residential")
```






## Police Safety 

```{r Police Safety}

forecast_chart(by_type_rt_forecast, "Code Zero")
```


## Minor Injuries 

```{r Minor Injuries}

forecast_chart(by_type_rt_forecast, "Collapse/Illness/Injury/Trapped")
```

## Complaints Against Police

```{r Compaints Against Police}

forecast_chart(by_type_rt_forecast, "Complaints Against Police")
```



## Concern for Safety 

```{r Concern For Safety}
forecast_chart(by_type_rt_forecast, "Concern For Safety")
```


## Criminal Damage 

```{r Criminal Damage}
forecast_chart(by_type_rt_forecast, "Criminal Damage")
```



## Domestic Incident

```{r Domestic Incident}
forecast_chart(by_type_rt_forecast, "Domestic Incident")
```




## Drug Related Offences 


```{r Drug Related Offences}
forecast_chart(by_type_rt_forecast,"Drugs")
```




## Firearms 


```{r Firearms}
forecast_chart(by_type_rt_forecast, "firearms")
```



## Fraud


```{r Fraud}
forecast_chart(by_type_rt_forecast,"Fraud")
```



## Hate Incident 


```{r Hate Incident}
forecast_chart(by_type_rt_forecast,"Hate Incident")
```



## Highway Disruption 

```{r Highway Disruption}
forecast_chart(by_type_rt_forecast,"Highway Disruption")
```




## Missing Person 

```{r Missing Person}
forecast_chart(by_type_rt_forecast, "Missing Person")
```



## Other - Admin Calls

```{r Other -Admin Calls}
forecast_chart(by_type_rt_forecast,"other - admin")
```


## Other - minor 


```{r Other - minor}
forecast_chart(by_type_rt_forecast, "other - minor")
```


 
## Police Generated 

```{r Police Generated}
forecast_chart(by_type_rt_forecast, "Police Generated")
```



## Potential Breach of the Peace


```{r Breach of Peace}
forecast_chart(by_type_rt_forecast,"Prevent Potential Breach of the Peace")
```



## Property Lost/Found 


```{r Property Lost/Found}
forecast_chart(by_type_rt_forecast,"Property Lost/Found")
```




## Road Related Traffic Offence

```{r Road Related Traffic}
forecast_chart(by_type_rt_forecast,"Road Related Traffic Offence")
```


## Robbery 

```{r Robbery}
forecast_chart(by_type_rt_forecast, "Robbery")
```


## Sexual Offences


```{r Sexual Offences}
forecast_chart(by_type_rt_forecast, "Sexual Offences")
```




## Shoplifitng

```{r Shoplifting}
forecast_chart(by_type_rt_forecast, "Shoplifting")
```



## Sudden Death 

```{r Sudden Death}
forecast_chart(by_type_rt_forecast, "Sudden Death")
```



## Suspicious Activity 

```{r Suspicious Activity}
forecast_chart(by_type_rt_forecast, "Suspicious Activity")
```



## Suspicious Package 

```{r Suspicious Package}
forecast_chart(by_type_rt_forecast, "Suspicious Package")
```



## Theft from Motor Vehicle

```{r Theft From Motor Vehicle}
forecast_chart(by_type_rt_forecast, "Theft From Motor Vehicle")
```



## Theft of Motor Vehicle 


```{r Theft of Motor Vehicle}
forecast_chart(by_type_rt_forecast, "Theft Of Motor Vehicle")
```



## Theft Other 

```{r Theft Other}
forecast_chart(by_type_rt_forecast, "Theft Other")
```



## Traffic collision

```{r Traffic Collisions}
forecast_chart(by_type_rt_forecast, "traffic collision")
```



## Violence and Harassment Incidents 

```{r Violence/Harassment}
forecast_chart(by_type_rt_forecast, "Violence/Harassment")
```







