---
title: "Shift patterns"
author: "RS MA NK EK"
date: "06/12/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(lubridate)
library(tsibble)
library(tidyverse)
# devtools::install_github('Ather-Energy/ggTimeSeries')
library(ggTimeSeries)
library(fmsb)

# reka path
calls <- read_csv("/Volumes/n8_covid/n8_data.csv.gz") %>% # USE CORRECT PATH FOR CLEAN DATA FILE ON YOUR COMPUTER
  janitor::clean_names()

# calls <- read_csv("Z:\\n8_data_v2.csv.gz") %>%  janitor::clean_names()

# create a "time of day" variable
calls$hour <- hour(ymd_hms(calls$incident_date_time))
calls$time <- case_when(calls$hour >= 5 & calls$hour < 11 ~ "Early", 
                        calls$hour >= 11 & calls$hour < 17 ~ "Daytime", 
                        calls$hour >= 17 & calls$hour < 23 ~ "Evening", 
                        calls$hour >= 23 | calls$hour < 5 ~ "Late", 
                        TRUE ~ NA_character_)
calls$time <- factor(calls$time, levels = c("Early","Daytime","Evening","Late"))
calls <- calls %>% 
  mutate(inc_date = floor_date(ymd_hms(incident_date_time), unit = "day"))

```

## Change in call times

We also wanted to explore whether there was any shift in the time where calls were being made. Either for day of the week, or hour of the day (and also how these map into shift patterns). 

Overall we find no real differences. 

## Day of the week


```{r dayofweekradar}

# try for radar chart for pre 2020
  
  dow_count_pre <- calls %>% 
    filter(inc_date < dmy('31-01-2020')) %>% 
    mutate(dow = wday(inc_date, label = TRUE), 
           yr_wk = yearweek(inc_date)) %>% 
    group_by(dow, yr_wk) %>% 
    count() %>% 
    group_by(dow) %>% 
    summarise(avg_calls = round(mean(n, na.rm = TRUE),0))
  
wide_dow_pre <- dow_count_pre %>% 
  pivot_wider(names_from = dow, values_from = avg_calls)

rownames(wide_dow_pre) <- "pre-pandemic"

# now for 2020

dow_count_2020 <- calls %>% 
  filter(inc_date >= dmy('31-01-2020')) %>% 
  mutate(dow = wday(inc_date, label = TRUE), 
         yr_wk = yearweek(inc_date)) %>% 
  group_by(dow, yr_wk) %>% 
  count() %>% 
  group_by(dow) %>% 
  summarise(avg_calls = round(mean(n, na.rm = TRUE),0))

wide_dow_2020 <- dow_count_2020 %>% 
  pivot_wider(names_from = dow, values_from = avg_calls)

rownames(wide_dow_2020) <- "2020"

# bind together
radar_df <- rbind(rep(max(dow_count_pre$avg_calls)+10, 7),
                      rep(min(dow_count_2020$avg_calls)-10, 7), 
                      wide_dow_pre, 
                      wide_dow_2020)

# add colours
line_cols <- c('#7570b3', '#1b9e77')

# make chart
radarchart(radar_df, pcol = line_cols, plwd = 4, plty = 1)
legend(x=0.8, y=1.4, legend = c("Pre-pandemic", "2020"), bty = "n", pch=20 , col=line_cols , text.col = "black", cex=1.2)


```


## Shift patterns

To look at shift patterns we coded calls made on or after 05:00 until before 11:00 as "Early", calls made on or after 11 until before 17:00 as "Daytime", calls made on or after 17:00 until before 23:00 as "Evening" and calls made on or after 23:00 until before 05:00 as "Late". 


```{r shiftradar}

calls$time <- factor(calls$time, levels = c("Early", "Late", "Evening", "Daytime"))


# try for radar chart for pre 2020

time_count_pre <- calls %>% 
  filter(inc_date < dmy('31-01-2020')) %>% 
  mutate(yr_wk = yearweek(inc_date)) %>% 
  group_by(time, yr_wk) %>% 
  count() %>% 
  group_by(time) %>% 
  summarise(avg_calls = round(mean(n, na.rm = TRUE),0))

wide_time_pre <- time_count_pre %>% 
  pivot_wider(names_from = time, values_from = avg_calls)

rownames(wide_time_pre) <- "pre-pandemic"

# now for 2020

time_count_2020 <- calls %>% 
  filter(inc_date >= dmy('31-01-2020')) %>% 
  mutate(yr_wk = yearweek(inc_date)) %>% 
  group_by(time, yr_wk) %>% 
  count() %>% 
  group_by(time) %>% 
  summarise(avg_calls = round(mean(n, na.rm = TRUE),0))

wide_time_2020 <- time_count_2020 %>% 
  pivot_wider(names_from = time, values_from = avg_calls)

rownames(wide_time_2020) <- "2020"

# bind together
radar_df <- rbind(rep(max(time_count_pre$avg_calls)+30, 7),
                  rep(min(time_count_2020$avg_calls)-10, 7), 
                  wide_time_pre, 
                  wide_time_2020)

# add colours
line_cols <- c('#7570b3', '#1b9e77')

# make chart
radarchart(radar_df, pcol = line_cols, plwd = 4, plty = 1)
legend(x=0.8, y=1.4, legend = c("Pre-pandemic", "2020"), bty = "n", pch=20 , col=line_cols , text.col = "black", cex=1.2)


```

## Hour of the day

Finally we can look more granularly at hourly calls, but still, we see no different patterns between calls made before 13/01/2020 and those made after, during the lockdown. 

```{r radarbyhour}

hour_count_pre <- calls %>% 
  filter(inc_date < dmy('31-01-2020')) %>% 
  mutate(yr_wk = yearweek(inc_date)) %>% 
  group_by(hour, yr_wk) %>% 
  count() %>% 
  group_by(hour) %>% 
  summarise(avg_calls = round(mean(n, na.rm = TRUE),0))

wide_hour_pre <- hour_count_pre %>% 
  pivot_wider(names_from = hour, values_from = avg_calls)

rownames(wide_hour_pre) <- "pre-pandemic"

# now for 2020

hour_count_2020 <- calls %>% 
  filter(inc_date >= dmy('31-01-2020')) %>% 
  mutate(yr_wk = yearweek(inc_date)) %>% 
  group_by(hour, yr_wk) %>% 
  count() %>% 
  group_by(hour) %>% 
  summarise(avg_calls = round(mean(n, na.rm = TRUE),0))

wide_hour_2020 <- hour_count_2020 %>% 
  pivot_wider(names_from = hour, values_from = avg_calls)

rownames(wide_hour_2020) <- "2020"

# bind together
radar_df <- rbind(rep(max(hour_count_pre$avg_calls)+10, 7),
                  rep(min(hour_count_2020$avg_calls)-10, 7), 
                  wide_hour_pre, 
                  wide_hour_2020)

# add colours
line_cols <- c('#7570b3', '#1b9e77')

# make chart
radarchart(radar_df[,c(1, 24:2)], pcol = line_cols, plwd = 4, plty = 1)
legend(x=0.8, y=1.4, legend = c("Pre-pandemic", "2020"), bty = "n", pch=20 , col=line_cols , text.col = "black", cex=1.2)



```
